---
title: "Homework 3: Leaflet Maps"
author: "Min Yan BEH (mbeh)"
date: "9/28/2018"
output: html_document
---

# Install Packages

You might need to first install the `Tigris` package, which will be used to load Census shapefiles for NYC Counties.

```{r include=False}
# install.packages("tigris")
```

# Setup: Import Packages

The package dependencies for this homework are listed below.

```{r setup, include=FALSE}
require(RColorBrewer)
require(rgdal)
require(leaflet)
require(leaflet.extras)
require(tigris)
options(tigris_use_cache = TRUE)
```

# Load Geospatial Datasets

## NYC County shapefiles (Polygons)

```{r}
# Define state & counties to focus on for this homework
state <- "New York"
county.names <- c("New York", "Kings", "Queens", "Bronx", "Richmond")

# Load & subset TIGER/Line shapefile data from Census Bureau
ny.counties.load <- counties(state, cb = TRUE)
ny.counties <- ny.counties.load[ny.counties.load$NAME %in% county.names,]

# Custom Palette for NYC County Polygons
factpal <- colorFactor(brewer.pal(n = 11, name ="Spectral"), county.names)
```

## NYC Bridge data (Lines)

```{r}
bridges.load <- readOGR("./data/nyc_bridges.shp", layer = "nyc_bridges")
```

Note: the above shapefile was cleaned from NYC roads dataset (119353 datapoints), to reduce loading times.
The code below details how the bridge data was extracted from the original NYC roads dataset.
*NB: The original roads dataset was >700MB in size and is not contained in the repository. See README.md for download instructions.*

```{r}
# roads.load <- readOGR("./data/geo_export_6090a65b-3a0c-49e0-b3b1-4953105ee436.shp",
#                       layer = "geo_export_6090a65b-3a0c-49e0-b3b1-4953105ee436", 
#                       GDAL1_integer64_policy = TRUE)
# roads.subset <- roads.load[roads.load@data$rw_type == 3, ]    # ID 3 represents bridges
# writeOGR(roads.subset, dsn="./data/", layer = "nyc_bridges", driver = "ESRI Shapefile")
```

## NYC Landmarks data (Points)

```{r}
landmarks.subset <- readOGR("./data/landmarks_subset.geojson") # 378 landmarks

# Custom Palette for Landmarks
landmark.types <- as.character(unique(landmarks.subset$type))
landmark.palette <- colorFactor(c("#3182bd", "#636363"), landmark.types)
```

Note: the above GeoJSON file is a subset of the original NYC landmarks dataset (47790 datapoints), to reduce loading times.
The code below details how the subset (scenic & interior landmarks) was retrieved and saved to the new geojson file.
*NB: The original landmarks dataset was >30MB and is not contained in the repository. See README.md for download instructions.*

```{r}
# landmarks <- readOGR("./data/landmarks.geojson")
# landmarks$type <- as.character(landmarks$lm_type)

##### Subset landmarks data to show only scenic & interior landmarks
# landmarks.subset <- landmarks[landmarks$type != "Historic District", ]
# landmarks.subset <- landmarks.subset[landmarks.subset$type != "Individual Landmark", ]
# writeOGR(landmarks.subset, "./data/landmarks_subset.geojson", layer = "landmarks", driver = "GeoJSON")
```


# Plot Leaflet Map 

Elements of the Leaflet map include:

* NYC counties as polygons
* NYC bridges as polylines
* NYC landmarks as circle markers (clustered)
* A legend for variable colored circle markers
* A layers control for basemaps and overlay groups

```{r}
leaflet() %>%
  # Basemaps
  addTiles(urlTemplate = "http://mt0.google.com/vt/lyrs=m&hl=en&x={x}&y={y}&z={z}&s=Ga", # Source: Class 10 app.R Example
           attribution = "Google", group="Google Theme") %>%
  addProviderTiles("CartoDB.DarkMatter", group = "Dark Theme") %>%
  addProviderTiles("Stamen.TonerLite", group = "B&W Theme") %>%
  
  # Overlay map: counties as polygons
  addPolygons(data = ny.counties, group = "Counties",
              fillColor = ~factpal(NAME), 
              fillOpacity = 0.4, 
              color = 'black', 
              weight = 2) %>%
  
  # Overlay map: bridges as polylines
  addPolylines(data = bridges.load, color = "#4a1486", weight = 3, opacity = 1, group = "Bridges") %>%
  
  # Overlay map: landmarks as points / circle markers
  addCircleMarkers(data = landmarks.subset, color = ~landmark.palette(type), # vary marker color based on landmark type
                   clusterOptions = markerClusterOptions(),            # cluster markers if they are densely populated
                   radius = 10, opacity = 1,                           # customize visual attributes
                   popup = ~paste0("<b>Street</b>: ", pluto_addr, "<br><b>Landmark</b>: ", lm_name), # define popup text
                   group = "Landmarks")
```